Non-contextual Hidden Variable Model for Quantum Mechanics

### 2017-12-23 version2postPRL/nonContextualv1.tex
Arvind Sir Update

\abstract{ \color{blue}
We construct a \sout{\color{red} simple} non-contextual
hidden variable model consistent with all {\color{red} the}
kinematic predictions of quantum mechanics (QM). The famous
Bell-KS theorem shows that non-contextual models which
satisfy a further reasonable restriction are inconsistent
with QM. {\color{red} In our construction,} we define a
weaker variant of this restriction which captures its
essence while still allowing a non-contextual description of
QM. This is in contrast to {\color{red} the} contextual
\sout{toy} hidden variable {\color{red} toy} models, such as
the one by Bell, {\color{red} and}  brings out an interesting
alternate way of looking at QM. {\color{red} The results}
also relate to {\color{red} the Bohmian}
model, where it is harder to pin
down such features.}
\begin{document}

@@

{\color{red} 
I think that we should use `functional consistency' with a
hyphen `functional-consistency' and avoid using the word
`functional inconsistency'. We should instead use 
`non functional-consistency'
}

\section{Introduction}
Quantum mechanics has been one of the most
successful theories in physics so far, however,
there has not yet been a final word on its
completeness and
interpretation~\cite{BellSpkblUnspkbl}.
Einstein's~\cite{EinsteinEPR} work on the
incompleteness of QM and the subsequent seminal
work of Bell~\cite{BellSpkblUnspkbl}, assessing
the compatibility of a more complete model
involving hidden variables (HV) and locality with
QM, has provided deep insights into  how the
quantum world differs from its classical
counterpart.  In recent times, these insights have
been of pragmatic utility in the area of quantum
information processing (QIP), where EPR pairs are
fundamental motifs of
entanglement~\cite{Ekert,PironioRndmnssCrtfcn,NielsenChuang}.
The work of  Kochen Specker (KS)~\cite{KochenSpecker}
and Gleason {\it
et.\,al.}~\cite{Gleason,Peres,Mermin} broadened
the schism between  HV models and QM.  They showed
that it was contextuality and not non-locality
which was at the heart of this schism and the
incompatibility between HV models and QM can arise
even for a single quantum system.  Contextuality
has thus been identified as a fundamental
non-classical feature of the quantum world and
experiments have also been proposed and conducted
to this effect~\cite{SimonContExpProp,
HuangContExp}.
Contextuality, on the one hand has
led to investigations on the foundational aspects
of QM including attempts to prove completeness of
QM~\cite{PawelCntxClsscl,CabelloMmryQM}, and on
the other hand has been harnessed for computation
and
cryptography~\cite{HowardCntxCmptn,CabelloCntxScrt}.
{\color{blue} While there have been generalisations, in this
letter, we restrict ourselves to the standard notion of
non-contextuality as used by KS~\cite{KochenSpecker}.}

Not all HV models (e.g.
Bohm's model based on
trajectories~\cite{Bohm1,Bohm2}), however, are incompatible
with QM~\cite{BellOnHiddenVariables} and we, in
this letter, present a new non-contextual HV model
consistent with QM.
\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{block2}
\caption{Non-contextuality is not inconsistent
with kinematic predictions of QM} \label{fig:block}\end{figure}
{\color{blue}
The KS theorem is applicable to non-contextual HV models
which additionally satisfy functional consistency --
{\it algebraic constraints obeyed by commuting quantum
observables must be satisfied by the HV model at the level
of individual outcomes}. 
One of the main justifications \sout{\color{red} given} for imposing
this was the \sout{reasonable} requirement that \sout{an}
{\color{red} the} observable $\hat A^2$ must depend on $\hat
A$ even at the HV level. This sounds reasonable because
otherwise one can construct HV models where these
observables {\color{red} can} get mapped to random variables
with no relation to each other (see~\cite{KochenSpecker}). 
As we
will show, one can weaken this requirement into what we call
weak functional consistency -- {\it functional consistency is
demanded only when the observables in question have sharp
values}. This entails that even in the HV model the
observables will depend on each other where they should for
{\color{red} consistency} and not otherwise,
thereby capturing the same idea {\color{red} without undue
constraints}. The consequence of
demanding only the weaker variant is that the KS theorem no
longer applies and non-contextual HV models can be
consistently constructed.

{\color{red}  Sentence below needs editing}
This is in stark contrast with
contextual HV models, such as the toy
model\footnote{introduced in connection with Gleason's
Theorem} by Bell~\cite{BellOnHiddenVariables}, which we
demonstrate by re-examining a `proof of contextuality' using
our model.}
\sout{
A closer scrutiny of the arguments
leading to the dichotomy between QM and HV models
reveals that the algebraic constraints obeyed by
quantum observables may not necessarily be obeyed
by the predictions of an HV model at the level of
individual outcomes and  that contextuality is 
not a necessary feature of QM. Those HV
models that obey such algebraic constraints
possess a special property which we call
``multiplicativity''.  Enlarging the scope of HV
models to include non-multiplicative models,
permits us to construct an explicit non-contextual
HV model consistent with 
kinematic predictions of
QM  %(see \figref{block}).
The re-examination of the very proof of
``contextuality''  is used to illustrate
non-multiplicativity, an intrinsic quantum
feature, which we propose as a possible
alternative to contextuality. 
Non-mulitiplicativity thus might turn out to be
the key quantum feature distinguishing the quantum
world from the classical one and may have a role
to play in QIP protocols. }
This provides a new
way to look at the classical-quantum divide and at
the foundations of quantum mechanics.

\section{Non-Contextuality and Functional Consistency}
We introduce some notation and make the relevant notions
precise to facilitate the description and assessment of our
model.
\begin{notn*}
(a) $\psi\in\mathcal{H}$ represents a pure quantum
mechanical state of the system in the Hilbert
space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is
defined to mean the set of Hermitian observables
for the system, (c) $[\mathcal{H}]$ is defined to
mean $\{\mathcal{H},\,\mathbb{R}^{\otimes}\}$, which
represents the state  space of the
system including HVs, (d) $[\psi]\in[\mathcal{H}]$
will represent the state of the system including
HVs, (e) a prediction map is $M:\{
\hat{\mathcal{H}},[\mathcal{H}] \}\to\mathbb{R}$,
(f) a sequence map is $S:\{
\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}
\}\to[\mathcal{H}]$, (g) $f$ is an arbitrary map
from $\{
\hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}}
\} \to \hat{\mathcal{H}}$ constructed using
multiplication and addition of compatible
observables, and multiplication with complex
numbers, (h) $\tilde{f}$ is a map constructed by
replacing observables in $f$ with real numbers.
\end{notn*}
\begin{defn} A theory is non-contextual, if it
provides a map $M: \{
\hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$
to explain measurement outcomes. A theory which is
not non-contextual is contextual~\cite{peresBook}.
\end{defn}

@@

\begin{defnrem}
{\color{blue} Broader definitions in the literature have
been suggested~\cite{spekkens_pra} which extend the notion
to probabilistic HV models. 
{\color{red} Sentence below is not clear. Do we
want to define measurement contextuality? or general
contextuality in this broader sense?}
The idea is that any feature of
a HV model that is not determined solely by the operational
aspect of QM is defined to be, in the case of measurements
for instance, measurement contextuality. 
To maintain a
distinction between different features of HV models, which
is of interest here, we stick to the standard definition.}

\sout{
Broader definitions in the literature have been
suggested which declare a larger set of theories
as non-contextual. For our %~\cite{spekkens_pra}. For our
purposes, however, this restricted definition will
suffice  and our results will in any case be valid
for the broader definitions. The term contextual
is used to suggest that the value an observable
takes might depend on which other compatible
observable it is being  measured with. }
\end{defnrem}

@@

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps, that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{functionally consistent} must be incompatible with
QM. 
This leaves open an interesting possibility that non-contextual maps which have \emph{weak functional consistency} could be consistent with QM.
{ \color{blue} Before proceeding to do so explicitly we observe that \emph{weak functional consistency} is, in fact, a consequence of QM.}
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{weak functional consistency} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined (in \defnref{seqnMltpl})
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{weak functional consistency}
must always hold.

@@

\section{Implications and Remarks}
{ \color{blue} 
The model demonstrates \emph{function inconsistency} as an alternative signature of quantumness as opposed to contextuality. This view is not just an artifact of the simplicity of the model. 
It }has implications to Bohm's HV
model (BHVM), where if the initial conditions are
precisely known the entire trajectory of a
particle (guided by the wavefunction) can be
predicted including the individual outcome of
measurements.  BHVM applied to a single spin-half
particle in a Stern-Gerlach experiment, can
predict opposite results for two measurements of
the same operator.  This observation which is
often used as a demonstration of ``contextuality''
in BM, does not involve  overlapping sets of
compatible measurements to provide two different
contexts~\cite{HardyCntxBM} {\color{blue}and is therefore not of
interest here.}  It turns out that,
if one constructs  a one-one map between an
experiment and an observable (by following a
certain convention), this so-called ``contextuality''
can be removed from BHVM.  However, the prediction
map so obtained from observables to measurement
outcomes turns out to be \emph{functionally inconsistent},
suggesting that \emph{functional inconsistency} is a more
suitable explanation of non-classicality
of BHVM.
{\color{blue} In fact, our model when appropriately extended
to continuous variables yields Bohmian trajectories in the
single particle  case which suggests that it can be used as
a starting point for constructing more interesting families
of HV theories that take quantum time dynamics into account
as well.

@@

{\color{blue} Second, note that} \sout{Clearly,} entanglement is not necessary to demonstrate
functional inconsistency; for instance the
Peres Mermin test is a state independent test
where a separable state can be used to arrive at a
contradiction. On the other hand if everything is
functionally consistent in a situation, can we have
violation of Bell's inequality or non-locality?
The answer is no and, therefore, one can say that
Bell's inequalities bring out  non-local
consequences of functionally inconsistent prediction
maps and the notion of functional inconsistency is
more basic.

@@

\acknowledgments
\sout{We are thankful to the anonymous referees for 
their feedback.} 
ASA and Arvind acknowledge
funding from KVPY and DST Grant No.
EMR/2014/000297 respectively.

>>>>

\abstract{ \color{blue}
We construct a simple non-contextual hidden variable model consistent with all kinematic predictions of quantum mechanics (QM). The famous Bell-KS theorem shows that non-contextual models which satisfy a further reasonable restriction are inconsistent with QM. We define a weaker variant of this restriction which captures its essence while still allowing a non-contextual description of QM. This is in contrast to contextual toy hidden variable models, such as the one by Bell, and brings out an interesting alternate way of looking at QM. This is also related to Bohm's model where it is harder to pin down such features.}

\begin{document}

@@

\section{Introduction}
Quantum mechanics has been one of the most
successful theories in physics so far, however,
there has not yet been a final word on its
completeness and
interpretation~\cite{BellSpkblUnspkbl}.
Einstein's~\cite{EinsteinEPR} work on the
incompleteness of QM and the subsequent seminal
work of Bell~\cite{BellSpkblUnspkbl}, assessing
the compatibility of a more complete model
involving hidden variables (HV) and locality with
QM, has provided deep insights into  how the
quantum world differs from its classical
counterpart.  In recent times, these insights have
been of pragmatic utility in the area of quantum
information processing (QIP), where EPR pairs are
fundamental motifs of
entanglement~\cite{Ekert,PironioRndmnssCrtfcn,NielsenChuang}.
The work of  Kochen Specker (KS)~\cite{KochenSpecker}
and Gleason {\it
et.\,al.}~\cite{Gleason,Peres,Mermin} broadened
the schism between  HV models and QM.  They showed
that it was contextuality and not non-locality
which was at the heart of this schism and the
incompatibility between HV models and QM can arise
even for a single quantum system.  Contextuality
has thus been identified as a fundamental
non-classical feature of the quantum world and
experiments have also been proposed and conducted
to this effect~\cite{SimonContExpProp,
HuangContExp}.
Contextuality, on the one hand has
led to investigations on the foundational aspects
of QM including attempts to prove completeness of
QM~\cite{PawelCntxClsscl,CabelloMmryQM}, and on
the other hand has been harnessed for computation
and
cryptography~\cite{HowardCntxCmptn,CabelloCntxScrt}.
{\color{blue} While there have been generalisations, in this letter, we restrict ourselves to the standard notion of non-contextuality as used by KS~\cite{KochenSpecker}.} 
Not all HV models (e.g.
Bohm's model based on
trajectories~\cite{Bohm1,Bohm2}), however, are incompatible
with QM~\cite{BellOnHiddenVariables} and we, in
this letter, present a new non-contextual HV model
consistent with QM.
\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{block2}
\caption{Non-contextuality is not inconsistent
with kinematic predictions of QM} \label{fig:block}\end{figure}
{\color{blue}
The KS theorem is applicable to non-contextual HV models which additionally satisfy functional consistency --- algebraic constraints obeyed by commuting quantum observables must be satisfied by the HV model at the level of individual outcomes. One of the main justifications given for imposing this was the reasonable requirement that an observable $\hat A^2$ must depend on $\hat A$ even at the HV level because otherwise one can construct HV models where these observables get mapped to random variables with no relation to each other (see~\cite{KochenSpecker}). As we will show, one can weaken this requirement into what we call weak functional consistency --- functional consistency is demanded only when the observables in question have sharp values. This entails that even in the HV model the observables will depend on each other where they should thereby capturing the same idea. The consequence of demanding only the weaker variant is that the KS theorem no longer applies and non-contextual HV models can be consistently constructed. This is in stark contrast with contextual HV models, such as the toy model\footnote{introduced in connection with Gleason's Theorem} by Bell~\cite{BellOnHiddenVariables}, which we demonstrate by re-examining a `proof of contextuality' using our model. 
}
\sout{
A closer scrutiny of the arguments
leading to the dichotomy between QM and HV models
reveals that the algebraic constraints obeyed by
quantum observables may not necessarily be obeyed
by the predictions of an HV model at the level of
individual outcomes and  that contextuality is 
not a necessary feature of QM. Those HV
models that obey such algebraic constraints
possess a special property which we call
``multiplicativity''.  Enlarging the scope of HV
models to include non-multiplicative models,
permits us to construct an explicit non-contextual
HV model consistent with 
kinematic predictions of
QM  %(see \figref{block}).
The re-examination of the very proof of
``contextuality''  is used to illustrate
non-multiplicativity, an intrinsic quantum
feature, which we propose as a possible
alternative to contextuality. 
Non-mulitiplicativity thus might turn out to be
the key quantum feature distinguishing the quantum
world from the classical one and may have a role
to play in QIP protocols. }
This provides a new
way to look at the classical-quantum divide and at
the foundations of quantum mechanics.

\section{Non-Contextuality and Functional Consistency}
We introduce some notation and make the relevant notions precise to facilitate the description and assesment of our model.
\begin{notn*}
(a) $\psi\in\mathcal{H}$ represents a pure quantum
mechanical state of the system in the Hilbert
space $\mathcal{H}$, (b) $\hat{\mathcal{H}}$ is
defined to mean the set of Hermitian observables
for the system, (c) $[\mathcal{H}]$ is defined to
mean $\{\mathcal{H},\,\mathbb{R}^{\otimes}\}$, which
represents the state  space of the
system including HVs, (d) $[\psi]\in[\mathcal{H}]$
will represent the state of the system including
HVs, (e) a prediction map is $M:\{
\hat{\mathcal{H}},[\mathcal{H}] \}\to\mathbb{R}$,
(f) a sequence map is $S:\{
\hat{\mathcal{H}},[\mathcal{H}],\mathbb{R}
\}\to[\mathcal{H}]$, (g) $f$ is an arbitrary map
from $\{
\hat{\mathcal{H}},\hat{\mathcal{H}},\dots\hat{\mathcal{H}}
\} \to \hat{\mathcal{H}}$ constructed using
multiplication and addition of compatible
observables, and multiplication with complex
numbers, (h) $\tilde{f}$ is a map constructed by
replacing observables in $f$ with real numbers.
\end{notn*}
\begin{defn} A theory is non-contextual, if it
provides a map $M: \{
\hat{\mathcal{H}},[\mathcal{H}] \} \to\mathbb{R}$
to explain measurement outcomes. A theory which is
not non-contextual is contextual~\cite{peresBook}.
\end{defn}

@@

\begin{defnrem}
{\color{blue} Broader definitions in the literature have been suggested~\cite{spekkens_pra} which extend the notion to probabilistic HV models. The idea is that any feature of a HV model that is not determined solely by the operational aspect of QM is defined to be, in the case of measurements for instance, measurement contextuality. To maintain a distinction between different features of HV models, which is of interest here, we stick to the standard definition.}
\sout{
Broader definitions in the literature have been
suggested which declare a larger set of theories
as non-contextual. For our %~\cite{spekkens_pra}. For our
purposes, however, this restricted definition will
suffice  and our results will in any case be valid
for the broader definitions. The term contextual
is used to suggest that the value an observable
takes might depend on which other compatible
observable it is being  measured with. }
\end{defnrem}

@@

\begin{rem} One could in principle assume $M$, to
be s.t. (a) $M(\hat{\mathbb{I}})=1$, (b)
$M(\alpha\hat{B}_{i})=\alpha M(\hat{B}_{i})$, for
$\alpha\in\mathbb{R}$, (c)
$M(\hat{B}_{i}^{2})=M(\hat{B}_{i})^{2}$, (d)
$M(\hat{B}_{i}+\hat{B}_{j})=M(\hat{B}_{i})+M(\hat{B}_{j})$,
to deduce (d)
$M(\hat{B}_{i}\hat{B}_{j})=M(\hat{B}_{i})M(\hat{B}_{j})$
and that $M(\hat{B}_{i})\in$ spectrum of
$\hat{B}_{i}$. Effectively then, condition (b)
listed in the theorem is satisfied as a
consequence.  Therefore, assuming (a)-(d) as
listed above, rules out a larger class of
$M$~\cite{KochenSpecker}.
\end{rem} 
Here $M$ maybe viewed as a specific class of
prediction maps, that implicitly depends on the
state $[\psi]$. It is clear that according to the
theorem, non-contextual maps which are
\emph{functionally consistent} must be incompatible with
QM. 
This leaves open an interesting possibility that non-contextual maps which have \emph{weak functional consistency} could be consistent with QM.
{ \color{blue} Before proceeding to do so explicitly we observe that \emph{weak functional consistency} is, in fact, a consequnce of QM.}
\begin{prop*} 
Let a quantum mechanical 
system be in a state, s.t. measurement of
$\hat{C}$ yields repeatable results (same result
each time). Then according to QM, \emph{weak functional consistency} holds, where $\hat{C}:=
f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n})$, and
$\hat{B}_{i}$ are as defined (in \defnref{seqnMltpl})
\end{prop*}
\begin{proof} 
Without loss of generality we can take
$\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}$ to be
mutually compatible and a complete set of
observables (operators can be added to make the set
complete if needed).
It follows
that $\exists$
$\left|\mathbf{b}=\left(b_{1},b_{2},\dots
b_{n}\right)\right\rangle $ s.t.
$\hat{B}_{i}\left|\mathbf{b}\right\rangle
=b_{i}\left|\mathbf{b}\right\rangle $,
and that
$\sum_{\mathbf{b}}\left|\mathbf{b}\right\rangle
\left\langle \mathbf{b}\right|=\hat{\mathbb{I}}$.
Let the state of the system
$\left|\psi\right\rangle $ be s.t.
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $. For
the statement to follow, one need only show that
$\left|\psi\right\rangle $ must be made of only
those $\left|\mathbf{b}\right\rangle $s, which
satisfy $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$.  This is the crucial step and
proving this is straightforward. We start with
$\hat{C}\left|\psi\right\rangle
=c\left|\psi\right\rangle $ and take its inner
product with $\left\langle \mathbf{b}\right|$ to
get 
\begin{eqnarray*} 
c \langle \mathbf{b}|\psi\rangle 
& = &
\langle
\mathbf{b} |\hat{C} |\psi
\rangle  
\\
&=&
\langle
\mathbf{b}| f(\hat{B}_{1},\hat{B}_{2},\dots\hat{B}_{n}) |
\psi \rangle
\\ 
&=&
\tilde f (b_{1},b_{2},\dots b_{n})
\langle
\mathbf{b}|\psi\rangle  
\end{eqnarray*}
Also, we have $\left|\psi\right\rangle
=\sum_{\mathbf{b}}\left\langle
\mathbf{b}|\psi\right\rangle
\left|\mathbf{b}\right\rangle $, from
completeness. If we consider
$\left|\mathbf{b}\right\rangle $s for which
$\left\langle \mathbf{b}|\psi\right\rangle \neq0$,
then we find that indeed $c=\tilde
f(b_{1},b_{2},\dots
b_{n})$. The case when $\left\langle
\mathbf{b}|\psi\right\rangle =0$ is anyway
irrelevant as the corresponding $\vert b\rangle$
does not contribute to $\vert \psi\rangle$.
We can thus conclude that
$\left|\psi\right\rangle $ is made only of those
$\left|\mathbf{b}\right\rangle $s that satisfy the
required relation.
\end{proof} 
It is worth noting that in the Peres Mermin
case, where $\hat{R}_{i}$ and $\hat{C}_{j}$ are
$\pm\hat{\mathbb{I}}$, it follows that all
states are their eigenstates. Consequently, for
these operators \emph{weak functional consistency}
must always hold.

@@

\section{Implications and Remarks}
{ \color{blue} 
The model demonstrates \emph{function inconsistency} as an alternative signature of quantumness as opposed to contextuality. This view is not just an artifact of the simplicity of the model. 
It }has implications to Bohm's HV
model (BHVM), where if the initial conditions are
precisely known the entire trajectory of a
particle (guided by the wavefunction) can be
predicted including the individual outcome of
measurements.  BHVM applied to a single spin-half
particle in a Stern-Gerlach experiment, can
predict opposite results for two measurements of
the same operator.  This observation which is
often used as a demonstration of ``contextuality''
in BM, does not inolve  overlapping sets of
compatible measurements to provide two different
contexts~\cite{HardyCntxBM} {\color{blue}and is therefore not of
interest here.}  It turns out that,
if one constructs  a one-one map between an
experiment and an observable (by following a
certain convention), this so-called ``contextuality''
can be removed from BHVM.  However, the prediction
map so obtained from observables to measurement
outcomes turns out to be \emph{functionally inconsistent},
suggesting that \emph{functional inconsistency} is a more
suitable explanation of non-classicality
of BHVM.
{\color{blue} In fact, our model when appropriately extended to continous variables yields Bohmian trajectories in the single particle  case which suggests that it can be used as a starting point for constructing more interesting families of HV theories that take quantum time dynamics into account as well.

@@

{\color{blue} Second, note that} \sout{Clearly,} entanglement is not necessary to demonstrate
functional inconsistency; for instance the
Peres Mermin test is a state independent test
where a separable state can be used to arrive at a
contradiction. On the other hand if everyting is
functionally consistent in a situation, can we have
violation of Bell's inequality or non-locality?
The answer is no and, therefore, one can say that
Bell's inequalities bring out  non-local
consequences of functionally inconsistent prediction
maps and the notion of functional inconsistency is
more basic.

@@

\acknowledgments
We are thankful to the anonymous referees for their feedback. ASA and Arvind acknowledge
funding from KVPY and DST Grant No.
EMR/2014/000297 respectively.
